Mallows, Black-box combinatorial optimization, limited budget, ???

### 2020-10-23 paper/main.tex
paper update
more comments

\begin{abstract}
%The abstract should briefly summarize the contents of the paper in
%150--250 words.
  Black-box combinatorial optimization problems arise in practice when the
  objective function is evaluated by means of a simulator or a real-world
  experiment. In such cases, classical techniques such as mixed-integer
  programming and local search cannot be applied. Moreover, often each solution
  evaluation is expensive in terms of time or resources, thus only a limited
  number of evaluations is possible, typically several order of magnitude
  smaller than in white-box optimization problems. In the continuous case,
  Bayesian optimization, in particular using Gaussian processes, has proven
  very effective under these conditions. Much less research is available in the
  combinatorial case. In this paper, we propose and analyze an
  estimation-of-distribution (EDA) algorithm based on a Mallows probabilistic
  model and compare it with CEGO, a Bayesian optimization algorithm for
  combinatorial optimization. Experimental results show that the Bayesian
  algorithm is able to obtain very good solutions with very few evaluations,
  however, at a significant computational cost, whereas the proposed EDA
  outperforms CEGO when the number of solutions evaluated approaches 400, and
  it is significantly faster. These results suggest that the combination of
  Bayesian optimization and a Mallows-based EDA may be an interesting direction
  for future research.
\textcolor{red}{16 pages max, deadline: 1 November 2020}
\keywords{Combinatorial optimization \and Bayesian optimization \and Expensive black-box optimization \and Estimation of distribution algorithms}
\end{abstract}

@@

In this work, we propose and analyze uMM, an EDA specifically designed for
black-box combinatorial optimization under a limited budget. The proposed uMM 
\ldots\MANUEL{what is the inspiration}
\EKHINE{we propose a probabilistic algrithm that (like cego?) (1) iteratively adds solutions to the sample and (2) fits a surrogate model for every new sample obtained. However,  different fom cego, our surrogate model is an estimation of the predicted fitness function at each point. CEGO controls the  intensification-explotation balance with the expected improvement??? We mimic the intensification-explotation balance by automatically controllong the variance of our surrogate model.}

@@

Let us assume a black-box ``fitness'' function that must be minimized over a
space of permutations $f\colon S_n \to \mathbb{R}$. Being black-box means that
we can evaluate any candidate permutation $\sigma \in S_n$ to obtain
$f(\sigma)$, however, we do not know anything else about the form of
$f$. Moreover, the evaluation of $f$ is expensive in computation time or
resources and, thus, limits the budget of candidate permutations that we can
evaluate to a relatively low number. Here, we will study budgets lower than
400. Such expensive black-box combinatorial problems arise in
manufacturing~\citep{FerAlvDiaIglEna2014ants}, where the evaluation function
may involve expensive simulations for which no closed-form mathematical
description is available.

@@

Combinatorial Efficient Global Optimization
(CEGO)~\citep{ZaeStoFriFisNauBar2014} is an extension of the well-known EGO
method~\citep{JonSchWel98go} to unconstrained black-box combinatorial
optimization problems. In EGO, Gaussian process models are used as a surrogate
of the landscape of the expensive original problem. An optimization method
searches for solutions in the surrogate model by optimizing the expected
improvement criterion, which balances the expected mean and variance of the
chosen solution. Once a solution is chosen, it is evaluated on the actual
objective function and the result is used to update the surrogate-model,
hopefully increasing its predictive power.

CEGO replaces the Euclidean distance measure, used by the surrogate model in
EGO, with a distance measure appropriate to combinatorial
landscapes~\citep{ZaeStoBar2014:ppsn}, such as the Kendall distance for
permutations~\citep{?}. In CEGO, the surrogate model is explored by a GA with
crossover and mutation operators appropriate for the particular combinatorial
problem. The original paper notes that coupling the GA with local search does
not improve the results significantly since the model is anyway an inexact
estimation of the original objective
function~\citep[p.~875]{ZaeStoFriFisNauBar2014}.

What else do we need to say?

\subsection{Distributions over permutations}
Distributions over permutations~\cite{critchlow91} are functions that assign a probability value  $\Prob(\sigma)\in[0,1]$ to each permutation $\sigma \in S_n$. One of the most popular distributions is the Mallows Model (MM), which is considered as an analogous to the Gaussian distribution for permutations. The MM defines the probability of each permutation $\sigma\in S_n$ as follows:
%
\begin{equation}
\Prob(\sigma)=\frac{\exp(-\theta d(\sigma, \sigma_0))}{\psi}
\end{equation}
%
with two parameters, $\theta$ and $\sigma_0$. Permutation $\sigma_0\S_n$ is a reference permutation and has the largest probability value, i.e., the mode of the distribution. The probability of every permutation $\sigma\in S_n$ decays exponentially as its distance $d(\sigma,\sigma_0)$ increases, and the dispersion parameter $\theta$  controls this decay.  The normalization constant $\psi$ can be easily computed for the Kendall's $\tau$ distance as well as for the Hamming, Cayley and Ulam distances~\cite{IruCalLoz2016permallows}. we assume that the distance $d(\sigma,\sigma_0)$ is the Kendall's $\tau$ distance, which counts the number of items ranked in different order in both permutations, \MANUEL{X Can we define briefly Kendall's tau?}

\begin{equation}
\begin{split}
d(\sigma _{1},\sigma _{2})=|\{(i,j):i<j,&(\sigma _{1}(i)<\sigma _{1}(j)\wedge \sigma _{2}(i)>\sigma _{2}(j))\vee \\
&(\sigma _{1}(i)>\sigma _{1}(j)\wedge \sigma _{2}(i)<\sigma _{2}(j))\}|
 \end{split}
 \label{eq:kendall_decomp}
\end{equation}

An alternative definition for the Kendall's-$\tau$ distance counts the number of adjacent swaps that have to be made in $\sigma_1^{-1}$ to convert it in $\sigma_2^{-1}$ and therefore it is sometimes called \ekhine{swap distance? como la llamas tú?}{}

One of the fundamental problems in statistics is to learn the parameters that best fit a given sample of data. We will use the standard maximum likelihood estimation (MLE) approach to fit the parameters of a MM for a given collection of permutations. 
The MLE process\MANUEL{X what learning process?} is divided in two stages: first, we estimate the central permutation of the distribution, $\hat{\sigma}_0$ and, second, compute the dispersion parameter, $\hat\theta$~\cite{IruCalLoz2016permallows}.

The exact MLE for $\hat\sigma_0$\MANUEL{X for $\sigma_0$ o for $\hat{\sigma}_0$?} is given by the well-known Kemeny ranking~\cite{DwoKumNao2001rank}. Unfortunately,  obtaining such ranking is NP-hard~\cite{DwoKumNao2001rank}. An alternative to the Kemeny ranking is the Borda ranking~\cite{AliMei2011kemeny}, which has several advantages: (i) it requires polynomial computational time; (ii) it  guaranties  high quality parameters as an estimator for the MM~\cite{CarProSha2013votes}; and (iii) in a general optimization perspective, Borda error is bounded~\cite{CopFleRur2010ordering}.

The Borda ranking of a sample of permutations $S \subset S_n$ is computed as follows. 
\ekhine{First, for each item $i$, compute its Borda score $B(i) =  \sum_{t\in S}  \sigma_t(i)$. Then, assign rank 1 to the item with the smallest Borda score, rank 2 to the item with the second smallest score and so on, i.e., order the items by Borda score increasingly.}
{First, for each position $i$, compute its Borda score $B(i) =  \sum_{t\in S}  \sigma_t(i)$. Then, assign label 1 to the position with the smallest Borda score, label 2 to the position with the second smallest score and so on, i.e., order the positions by Borda score increasingly.}
\MANUEL{sorry, I don't follow here, what is an item? What does $\sigma_t(i)$ means? EKHINE: he cambiado las definiciones. $\sigma(i)=j$ es  $j$ es el ranking del item $i$}

\subsection{Unbalanced Borda}\label{sec:uborda}
We introduce now the core for our probabilistic algorithm, the uBorda algorithm. Following the convention of the community, the ranking returned by the uBorda algorithm will be denoted uBorda ranking. The uBorda algorithm is a recent generalization of Borda~\cite{???}.\MANUEL{X This paper is missing \ekhine{}{to be pocessed by arxiv}} In uBorda, we consider that each permutation in the sample $\sigma\in S$ has an associated weight $w(\sigma)$ where $w\colon S_n \to \mathbb{R_+}$. Then, the score of item $i$ is defined as follows:
%
\begin{equation}
\begin{split}
B(i) =  \sum_{t\in S}  w(\sigma_t) \sigma_t(i),
\end{split}
\label{eq:uborda_score}
\end{equation}
%
and, the same as Borda, orders the \ekhine{positions}{items} increasingly by their score.\MANUEL{X what is $w()$?}

Borda and uBorda are equivalent when $w(\sigma)$ is constant for every $\sigma$. However, in other case uBorda ranking will be closer to the rankings in $S$ with \ekhine{lower}{smaller} value of $w(\sigma)$. Rougly speaking, it is equivalent to replicating in the sample the $\sigma$ with low values of $w(\sigma$).
%Intuitively, the uBorda ranking a given sample of permutations $S$\MANUEL{X What is $S$?} is equivalent to the Borda ranking of $S'$, being $S'$ an extension of $S$ where each each $\sigma\in S$ has been replicated proportionally to its weight $w(\sigma)$.\MANUEL{X so $w$ is a weight function from $w\colon S_n \to \mathbb{R}$?} %
Clearly, the choice of  $w(\sigma)$ \MANUEL{X by weight you mean $\rho$?} depends on the relevant property of a ranking in a particular domain. For example, uBorda has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift~\cite{}.\MANUEL{X where? Do we miss a citation here?\ekhine{}{to be processed, arxiv}}

In this paper, we would like to replicate each permutation $\sigma$ a number of times that is inversely proportional to $f(\sigma)$.\MANUEL{We have not defined $f$ but I guess it is fintess, however, we minimize $f$, so how you handle that? X I guess we need to define the black-box optimization problem earlier than here} Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$. Interestingly, uBorda is computed in polynomial time. \MANUEL{by ``with smaller value'' do you mean ranking for increasing values of $f(\sigma)$?}

\section{Unbalanced Mallows model: probabilistic BB optimization}

@@

\ekhine{
cosas de intro : 
This new methodology relies on the assumption that the fitness landscatpe is correlated with a distribution over the same space. 
Previous paper have confirmed this point
For this case study, we show that this case for the Mallows model under the Kendall distance and 
In this section, we introduce our main contribution: We present a new methodology for the optimization of black box functions with a minimal number of queries. This methodology is a probabilistic, population-based algorithm. 
Essentially, the algorithm  considers a sample of permutations along with the fitness of each of them. It is an iterative process in which the number of iterations is fixed. 
Then, at each iteration, UMM (i) estimates a surrogate distribution of the fitness function and (ii) proposes a good fitted individual.}{aqui ha habido mas magia de git, esta seccion empieza en el siguiere paraafo} \MANUEL{Hay un poco de duplicacion entre lo anterior y lo siguiente, no?}

In this section we describe our main contribution: a novel algorithm for the optimization of a black box function over permutation spaces. This algorithm is called Unbalanced Mallows model (UMM) since it a probabilistic method based on a well-known probability distribution for permutations: the Mallows model.\MANUEL{this sounds as if this was the first time you talk about MM, but you already described it above, no? }

\begin{algorithm}[t]
 \caption{UMM: Unbalanced Mallows Model for Black-Box Optimization}
 \label{alg:umm}
  \begin{algorithmic}[1]
    \Require $n$: ?, $m$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
    \For{$i \assign 1$ \textbf{to} $m$ \text{evaluations}}
    \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$
    \State linear decrease on $\theta$
    \State $\sigma \assign$ sample from a $MM(\hat\sigma_0, \theta)$
    \State $S \assign S \cup \{\sigma\}$
    \State $F \assign F \cup \{f(\sigma)\}$
    \EndFor 
    \State \Return $S$, $F$
  \end{algorithmic}
\end{algorithm}

As shown in Algorithm~\ref{alg:umm}, UMM  starts with a \emph{small} sample of $m'$ permutations,\MANUEL{but the algorithm requires $m$, no $m'$?} $S=\{\sigma_1, \dotsc,  \sigma_{m'}\}$ that are randomly generated. UMM also keeps the function evaluation of each solution, $F=\{f(\sigma_1), \ldots, f(\sigma_{m'})\}$.  Then, for $m$ iterations\EKHINE{esto es exactamente así o es $m-m'$?}, UMM generates a random, promising candidate solution $\sigma$ from a probabilistic model and includes it in the sample $S$ while $f(\sigma)$ is included in $F$. At each iteration, the internal probabilistic model is updated.

The probabilistic model is a MM that assigns higher probability values to the permutations in which the fitness value is known or expected to be minimized. The main contribution of our approach is that the distribution does not represent the current sample $S$ considered by the algorithm but instead, represents a variation of it. This variation is based on the uBorda algorithm described in Section~\ref{sec:uBorda} and is abstractly equivalent to replicating in the sample the permutations \manuel{inversely proportionally}{a number of times inversely proportional} to their fitness values, i.e., there will be many copies of the good solutions and few of the bad ones. One key property of UMM is the computational complexity of working with the original sample and the transformed sample is the same, i.e., polynomial. However, the question of how many times should we replicate each permutation in the sample? This is controlled with the $\rho$ parameter in uBorda and its tunning and will be detailed in Section~\ref{sec:learning} and the experimental part.\MANUEL{X is this $\rho$ the same we explained earlier?}

At each iteration of the algorithm the following steps take place: (1) randomly
generate a permutation distributed according to the internal model and include
it in the sample and (2) evaluate the new permutation and update the
probabilistic model. These steps correspond to two statistical operations: (1)
sample a probability distribution and (2) learn a probability distribution of
the transformation of the sample. Both are describe in the subsequent sections.

\subsubsection{Learning step.}\label{sec:learning}
As shown in Section~\ref{sec:backgroud}, MM can be approximately learnt in two steps (1) estimating the location parameter $\sigma_0$ and (2) computing the scale parameter $\theta$. Since the scale parameter $\theta$ is set in the sampling step,\MANUEL{X How is it set automatically?} we focus on estimating $\sigma_0$.

The location parameter of a MM is usually approximated with the Borda algorithm. Since we are interested on biasing the model towards the regions of the space where the fitness function is smaller, we would like to replicate the permutations a number of times that is inversely proportionally to their fitness values.
We can simulate the replication of the best fitted individuals by using uBorda and setting $w(\sigma)=\rho^{f(\sigma)}$, as described in Section~\ref{sec:uborda}. This generalization requires setting a value for parameter $\rho$. We show how to choose this value in the experimental section.\MANUEL{I thought we select it automatically now?\ekhine{that is $\theta$}}

\subsubsection{Sampling step.}
Sampling can be done efficiently~\cite{IruCalLoz2016permallows} given the parameters that describe the model, i.e., $\sigma_0$ and $\theta$. Given these two parameters, sampling a MM has been studied in \cite{Collas}\cite{IruCalLoz2016permallows}\cite{Regenwetter2004}\MANUEL{You call this Regenwetter, but the author is Doignon?} where the computational complexities are $O(n\log n)$ and $O(n^2)$.
%
At each iteration of UMM, $\sigma_0$ is the permutation obtained as a solution in the previous learning step. The scale $\theta$, in the other hand, is set automatically by the algorithm to control the diversification and intensification trade off: At the first iteration it is set so that the expected distance equals  ${n\choose 4}$ (half of the distance of the uniform distribution), and then it linearly decrease until the last iteration where the expected distance equals 1. Similar approaches consisting of changing the variance to control the exploration-explotation tradeoff have been taken in different contexts~\cite{arza2019approaching}.\MANUEL{there is something wrong with this reference}
%
This setting reflects the fact that, despite believing that the best global optimum at each iteration is $\sigma_0$, the uncertainty at the first stages of the algorithm is high and\MANUEL{and?}

\subsection{Comparison to standard probabilistic optimization approaches.}
UMM is a population based algorithm. There exists a wide body of research on population based probabilistic algorithms, where most prominent examples are Estimation of Distribution Algorithms (EDAs). Our approach drastically differs from EDAs since EDAs assume that function evaluations are cheap and they can do many of them. On the contrary, UMM minimizes the number of function evaluations. Moreover, UMM does not rely on an evolving sequence of samples (called generations in evolutionary approaches). Instead, UMM considers a single sample of permutations during the whole execution that is increased iteratively, with one permutation each iteration.\MANUEL{I would suggest to move this subsection to the intro, is that OK?} 
\ekhine{bayesian optim}{ff}
\EKHINE{subir a arxiv uBorda}

@@

Complete this with details of uMM: MANUEL

>>>>

\begin{abstract}
%The abstract should briefly summarize the contents of the paper in
%150--250 words.
  Black-box combinatorial optimization problems arise in practice when the
  objective function is evaluated by means of a simulator or a real-world
  experiment. In such cases, classical techniques such as mixed-integer
  programming and local search cannot be applied. Moreover, often each solution
  evaluation is expensive in terms of time or resources, thus only a limited
  number of evaluations is possible, typically several order of magnitude
  smaller than in white-box optimization problems. In the continuous case,
  Bayesian optimization, in particular using Gaussian processes, has proven
  very effective under these conditions. Much less research is available in the
  combinatorial case. In this paper, we propose and analyze an
  estimation-of-distribution (EDA) algorithm based on a Mallows probabilistic
  model and compare it with CEGO, a Bayesian optimization algorithm for
  combinatorial optimization. Experimental results show that the Bayesian
  algorithm is able to obtain very good solutions with very few evaluations,
  however, at a significant computational cost, whereas the proposed EDA
  outperforms CEGO when the number of solutions evaluated approaches 400, and
  it is significantly faster. These results suggest that the combination of
  Bayesian optimization and a Mallows-based EDA may be an interesting direction
  for future research.\sloppy  
\textcolor{red}{16 pages max, deadline: 1 November 2020}
\keywords{Combinatorial optimization \and Bayesian optimization \and Expensive black-box optimization \and Estimation of distribution algorithms}
\end{abstract}

@@

In this work, we propose and analyze UMM, an EDA specifically designed for
black-box combinatorial optimization under a limited budget. The proposed UMM 
is a population based algorithm. There exists a wide body of research on population based probabilistic algorithms, where most prominent examples are Estimation of Distribution Algorithms (EDAs). Our approach drastically differs from EDAs since EDAs assume that function evaluations are cheap and they can do many of them. On the contrary, UMM minimizes the number of function evaluations. Moreover, UMM does not rely on an evolving sequence of samples (called generations in evolutionary approaches). Instead, UMM considers a single sample of permutations during the whole execution that is increased iteratively, with one permutation each iteration.

\ldots\MANUEL{what is the inspiration}
\EKHINE{we propose a probabilistic algrithm that (like cego?) (1) iteratively adds solutions to the sample and (2) fits a surrogate model for every new sample obtained. However,  different fom cego, our surrogate model is an estimation of the predicted fitness function at each point. CEGO controls the  intensification-explotation balance with the expected improvement??? We mimic the intensification-explotation balance by automatically controllong the variance of our surrogate model.}

@@

Let us assume a black-box ``fitness'' function that must be minimized over a
space of permutations $f\colon S_n \to \mathbb{R}$. Being black-box means that
we can evaluate any candidate permutation $\sigma \in S_n$ to obtain
$f(\sigma)$, however, we do not know anything else about the form of
$f$. Moreover, the evaluation of $f$ is expensive in computation time or
resources and, thus, limits the budget of candidate permutations that we can
evaluate to a relatively low number. Here, we will study budgets lower than
400. Such expensive black-box combinatorial problems arise, for example, in
manufacturing~\citep{FerAlvDiaIglEna2014ants}, where the fitness function
may involve expensive simulations for which no closed-form mathematical
description is available.

Distances in permutation space can be measured in various ways.  Kendall's $\tau$ distance counts the number of items ranked in different order in both permutations:
%
\begin{equation}
\begin{split}
d(\sigma _{1},\sigma_{2})=\left\lvert\Bigl\{(i,j):i<j \land \phantom{}\Bigl(\right.&\bigl(\sigma_{1}(i) < \sigma_{1}(j)\land \sigma_{2}(i)>\sigma_{2}(j)\bigr)\lor \phantom{} \\
 &\bigl(\sigma_{1}(i)>\sigma_{1}(j)\land \sigma_{2}(i)<\sigma_{2}(j)\bigr)\Bigr)\Bigr\}\Bigr\rvert\\
 \end{split}
 \label{eq:kendall_decomp}
\end{equation}
%
An alternative\MANUEL{Alternative or equivalent? That is, it is the same numbers or it will produce a different result?} definition for the Kendall's-$\tau$ distance counts the number of adjacent swaps that have to be made in $\sigma_1^{-1}$ to obtain $\sigma_2^{-1}$ and, therefore, it is sometimes called \emph{swap distance}~\citep{ZaeStoBar2014:ppsn}.

@@

Combinatorial Efficient Global Optimization
(CEGO)~\citep{ZaeStoFriFisNauBar2014} is an extension of the well-known EGO
method~\citep{JonSchWel98go} to unconstrained black-box combinatorial
optimization problems. In EGO, Gaussian process models are used as a surrogate
of the landscape of the expensive original problem. An optimization method
searches for solutions in the surrogate model by optimizing the expected
improvement criterion, which balances the expected mean and variance of the
chosen solution. Once a solution is chosen, it is evaluated on the actual
fitness function and the result is used to update the surrogate-model,
hopefully increasing its predictive power.

CEGO replaces the Euclidean distance measure, used by the surrogate model in
EGO, with a distance measure appropriate to combinatorial
landscapes~\citep{ZaeStoBar2014:ppsn}, such as Kendall's $\tau$ distance for
permutations (Eq.~\ref{eq:kendall_decomp}). In CEGO, the surrogate model is explored by a GA with
crossover and mutation operators appropriate for the particular combinatorial
problem. The original paper notes that coupling the GA with local search does
not improve the results significantly since the model is anyway an inexact
estimation of the original fitness
function~\citep[p.~875]{ZaeStoFriFisNauBar2014}.

\subsection{Distributions over permutations}
Distributions over permutations~\cite{critchlow91} are functions that assign a probability value  $\Prob(\sigma)\in[0,1]$ to each permutation $\sigma \in S_n$. One of the most popular distributions is the Mallows Model (MM), which is considered as an analogous to the Gaussian distribution for permutations. The MM defines the probability of each permutation $\sigma\in S_n$ as follows:\MANUEL{Could you check this? I wanted to introduce the notation $MM(\sigma_0, \theta)$ that we use later}
%
\begin{equation}\label{eq:MM}
\Prob(\sigma \mid \sigma_0, \theta )= \frac{\exp(-\theta d(\sigma, \sigma_0))}{\psi} \sim MM(\sigma_0, \theta)
\end{equation}
%
with two parameters, $\theta$ and $\sigma_0$. Permutation $\sigma_0\in S_n$ is a reference permutation and has the largest probability value, i.e., the mode of the distribution. The probability of every permutation $\sigma\in S_n$ decays exponentially as its distance $d(\sigma,\sigma_0)$ increases, and the dispersion parameter $\theta$  controls this decay.  The normalization constant $\psi$ can be easily computed for the Kendall's $\tau$ distance (Eq.~\ref{eq:kendall_decomp}) as well as for the Hamming, Cayley and Ulam distances~\cite{IruCalLoz2016permallows}.

\manuel{One of the fundamental problems in statistics is to learn the parameters that best fit a given sample of data.}{}\MANUEL{I think we don't need to emphasize this, it is known} %
We will use the standard maximum likelihood estimation (MLE) approach to fit the parameters of a MM for a given collection of permutations. 
The MLE process is divided in two stages: first, we estimate the central permutation of the distribution, $\hat{\sigma}_0$ and, second, compute the dispersion parameter, $\hat\theta$~\cite{IruCalLoz2016permallows}.

The exact MLE for $\hat\sigma_0$\MANUEL{MLE for $\sigma_0$ or for $\hat{\sigma}_0$?} is given by the well-known Kemeny ranking~\cite{DwoKumNao2001rank}. Unfortunately,  obtaining such ranking is NP-hard~\cite{DwoKumNao2001rank}. An alternative to the Kemeny ranking is the Borda ranking~\cite{AliMei2011kemeny}, which has several advantages: (i) it requires polynomial computational time; (ii) it  guaranties  high quality parameters as an estimator for the MM~\cite{CarProSha2013votes}; and (iii) in a general optimization perspective, Borda error is bounded~\cite{CopFleRur2010ordering}.\MANUEL{I'm not sure what the latter means}

The Borda ranking of a sample of permutations $S \subset S_n$ is computed as follows. 
First, for each position $i \in [n]$, compute its Borda score $B(i) =  \sum_{t\in S}  \sigma_t(i)$. Then, assign label 1 to the position with the smallest Borda score, label 2 to the position with the second smallest score and so on, i.e., order the positions by Borda score increasingly.

\subsection{Unbalanced Borda}\label{sec:uborda}
We introduce now the core for our probabilistic algorithm, the uBorda algorithm. Following the convention of the community, the ranking returned by the uBorda algorithm will be denoted uBorda ranking. The uBorda algorithm is a recent generalization of Borda~\cite{???}.\MANUEL{X This paper is missing \ekhine{}{to be pocessed by arxiv}} In uBorda, we consider that each permutation in the sample $\sigma\in S$ has an associated weight $w(\sigma)$ where $w\colon S_n \to \mathbb{R^{+}}$. Then, the score of \manuel{item}{position} $i$ is defined as follows:
%
\begin{equation}\label{eq:uborda_score}
B(i) =  \sum_{t\in S}  w(\sigma_t) \sigma_t(i) \enspace,
\end{equation}
%
and, the same as Borda, orders the \ekhine{positions}{items}\MANUEL{why items? we talk about positions above. We could talk about decision variables if that is clearer} increasingly by their score.

Borda and uBorda are equivalent when $w(\sigma)$ is constant for every $\sigma$. However, in other case uBorda ranking will be closer to the rankings in $S$ with smaller value of $w(\sigma)$. Roughly speaking, it is equivalent to replicating in the sample the $\sigma$ with low values of $w(\sigma$).
%Intuitively, the uBorda ranking a given sample of permutations $S$\MANUEL{X What is $S$?} is equivalent to the Borda ranking of $S'$, being $S'$ an extension of $S$ where each each $\sigma\in S$ has been replicated proportionally to its weight $w(\sigma)$.\MANUEL{X so $w$ is a weight function from $w\colon S_n \to \mathbb{R}$?} %
Clearly, the choice of  $w(\sigma)$ depends on the relevant property of a ranking in a particular domain. For example, uBorda has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift~\cite{}.\MANUEL{X where? Do we miss a citation here?\ekhine{}{to be processed, arxiv}}

In this paper, we would like to replicate each permutation $\sigma$ a number of times that is inversely proportional to $f(\sigma)$. Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$. Interestingly, uBorda is computed in polynomial time.

\renewcommand{\subsubsection}{\smallskip{}\noindent\textbf}

\section{Unbalanced Mallows model (UMM)}

% \ekhine{
% cosas de intro : 
% This new methodology relies on the assumption that the fitness landscatpe is correlated with a distribution over the same space. 
% Previous paper have confirmed this point
% For this case study, we show that this case for the Mallows model under the Kendall distance and 
% In this section, we introduce our main contribution: We present a new methodology for the optimization of black box functions with a minimal number of queries. This methodology is a probabilistic, population-based algorithm. 
% Essentially, the algorithm  considers a sample of permutations along with the fitness of each of them. It is an iterative process in which the number of iterations is fixed. 
% Then, at each iteration, UMM (i) estimates a surrogate distribution of the fitness function and (ii) proposes a good fitted individual.}{aqui ha habido mas magia de git, esta seccion empieza en el siguiere paraafo} \MANUEL{Hay un poco de duplicacion entre lo anterior y lo siguiente, no?}

In this section we describe our main contribution: Unbalanced Mallows model
(UMM), a novel algorithm for the optimization of a black-box function over
permutation spaces. This algorithm is a probabilistic method based on the
Mallows model and uBorda learning.

\begin{algorithm}[t]
 \caption{UMM: Unbalanced Mallows Model for Black-Box Optimization}
 \label{alg:umm}
  \begin{algorithmic}[1]
    \Require $\minit$: number of initial permutations, $m$: total budget, $\rho$: learning rate
    %\\ ?, $m$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
    \State $S \assign $ generate $\minit$ permutations uniformly at random
    \State $F \assign \{ f(\sigma) \mid \sigma \in S\}$ \Comment{Evaluate them}
    \For{$i \assign \minit$ \textbf{to} $m$ \text{evaluations}}
    \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$\label{line:uborda}
    \State linear decrease on $\theta$
    \State $\sigma \assign$ sampled from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
    \State $S \assign S \cup \{\sigma\}$, $\quad F \assign F \cup \{f(\sigma)\}$
    \EndFor 
    \State \Return $S$, $F$
  %   \Require $n$: ?, $m$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
  %   \For{$i \assign 1$ \textbf{to} $m$ \text{evaluations}}
  %   \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$
  %   \State linear decrease on $\theta$
  %   \State $\sigma \assign$ sample from a $MM(\hat\sigma_0, \theta)$
  %   \State $S \assign S \cup \{\sigma\}$
  %   \State $F \assign F \cup \{f(\sigma)\}$
  %   \EndFor 
  %   \State \Return $S$, $F$
 \end{algorithmic}
\end{algorithm}

As shown in Algorithm~\ref{alg:umm}, UMM  starts with a \emph{small} sample of $\minit$ permutations that are randomly generated (although it would be possible to employ Latin Hyper-Cube sampling or other methods) and evaluated. Then, using the current sample $S$, the uBorda algorithm is applied to find a reference permutation $\hat{\sigma}_0$ (line~\ref{line:uborda}). Together with the parameter $\theta$, they define a Mallows model (Eq.~\ref{eq:MM}) from which we sample a new candidate permutation (line~\ref{line:sample}). This candidate permutation (solution) is evaluated and added to the sample $S$. The process is repeated up to a given maximum budget of evaluations $m$.\MANUEL{We do not explain the initial value of $\theta$ or how we decrease it. Can we give a bit more detail?}

The key idea behind UMM is that the probabilistic Mallows model (MM) used in
UMM does not simply represent the current sample $S$. Instead, the model
assigns higher probability values to the permutations with known (or expected)
minimal fitness values. This is achieved by means of the uBorda ranking
(Section~\ref{sec:uBorda}), which can be seen as replicating each permutation
$\sigma$ in the sample a number of times inversely proportional to its fitness
$f(\sigma)$, i.e., it behaves as if there were many copies of the best
solutions and few of the bad ones. A crucial property of UMM is that the
computational complexity of working with the original sample and with the
transformed sample is the same, i.e., polynomial. The most important steps are
the sampling of a new permutation from $MM(\sigma_0,\rho)$ and the update
(\emph{learning}) of the parameters $\sigma_0$ and $\rho$.\MANUEL{Is
  this $\rho$ the same we explained earlier for the weights?}

% 'However, the question of how
% many times should we replicate each permutation in the sample? This is
% controlled with the $\rho$ parameter in uBorda and its tunning and will be
% detailed in Section~\ref{sec:learning} and the experimental part.

\manuel{At each iteration of the algorithm the following steps take place: (1) randomly
generate a permutation distributed according to the internal model and include
it in the sample and (2) evaluate the new permutation and update the
probabilistic model. These steps correspond to two statistical operations: (1)
sample a probability distribution and (2) learn a probability distribution of
the transformation of the sample. Both are describe in the subsequent sections.}{}\MANUEL{I don't think we need this, it is said above (if some detail is missing we should integrate it the paragraph above)}

\newcommand{\Edist}{\ensuremath{\text{E}[D]}}

\subsubsection{Sampling step.}
Sampling from $MM(\sigma_0,\rho)$  can be done efficiently~\cite{IruCalLoz2016permallows} given the parameters  $\sigma_0$ and $\theta$, as studied by \citet{Collas,IruCalLoz2016permallows,Regenwetter2004}\MANUEL{Which paper is Collas? It is not found.}\MANUEL{You call this Regenwetter, but the author is Doignon? \ekhine{}{he is another coauthor} But the entry shows only 1 author.} where the computational complexities are $O(n\log n)$ and $O(n^2)$.\MANUEL{respectively?}
%
At each iteration of UMM, $\sigma_0=\hat\sigma_0$ is the permutation obtained as a solution in the previous learning step. The scale $\theta$ controls the variance and expected distance $\Edist$ of the MM.\MANUEL{I'm not sure what is $D$, distance from where to where?} It is set automatically by the algorithm to control the diversification and intensification trade off. At the first iteration, it is set so that the expected distance $\Edist$ equals  $\binom{n}{4}$ (half of the distance of the uniform distribution), and then the expected distance linearly decrease until the last iteration where the expected distance equals 1. The expected distance, shown in the next line, does not have a closed-form as function of $\theta$ but its monotonicity guarantees low-cost, high-quality approximations:\MANUEL{I don't understand this sentence, sorry. You say it does not have a closed-form, but you show a closed-form. Do you mean that the formula below is only an approximation of $\Edist$? What is low-cost here? what is high-quality?}\MANUEL{What is $k$ in the equation?}
%
\begin{equation}\label{eq:expectation}
\Edist = \frac{k \cdot \exp(-\theta)}{1-\exp(-\theta)} - \sum_{j=n-k+1}^{n}  \frac{j \exp(-j \theta)}{1-\exp(-j\theta)}
\end{equation}

A purely explotation approach\ekhine{esto se dice asi? uno greedy igual esta mejor dicho} would systematically generate $\sigma_0$ in the sampling stage.\MANUEL{I'm not sure what ``systematically generate $\sigma_0$'' means.} However, this setting\MANUEL{which setting? $\theta$?}  controls the exploration-explotation tradeoff as a function of the number of function evaluations that the algorithm performs.  Similar approaches have been taken in different contexts~\cite{arza2019approaching}.\MANUEL{there is something wrong with this reference\ekhine{}{qué le pasa?} Don't you see a complete mess of the authors list? It looks like it has 20 authors.}

\subsubsection{Learning step.}
As explained above, the parameter $\theta$ is updated deterministically at the
sampling step. The parameter $\sigma_0$, however, is learned from the current
sample $S$ by using the uBorda algorithm with a weight
$w(\sigma)=\rho^{f(\sigma)}$ (Section~\ref{sec:uborda}). The learning rate
is controlled by parameter $\rho \in (0,1)$, which currently is user-defined. Higher values or $\rho$ bias the sampling towards the best solutions found so far, which may lead to quick stagnation, whereas smaller values make the sampling more uniform, thus slowing down the learning.

% As shown in Section~\ref{sec:backgroud}, MM can be approximately learnt in two steps (1) estimating the location parameter $\sigma_0$ and (2) computing the scale parameter $\theta$. Since the scale parameter $\theta$ is set in the sampling step,\MANUEL{X How is it set?} we focus on estimating $\sigma_0$.

% The location parameter of a MM is usually approximated with the Borda algorithm. Since we are interested on biasing the model towards the regions of the space where the fitness function is smaller, we would like to replicate the permutations a number of times that is inversely proportionally to their fitness values.
% We can simulate the replication of the best fitted individuals by using uBorda and setting $w(\sigma)=\rho^{f(\sigma)}$, as described in . This generalization requires setting a value for parameter $\rho$. We show how to choose this value in the experimental section.\MANUEL{I thought we select it automatically now?\ekhine{that is $\theta$}}

\ekhine{bayesian optim}{ff}
\EKHINE{subir a arxiv uBorda}

@@

@@

Complete this with details of UMM: MANUEL
